---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# covid19swiss

<!-- badges: start -->
[![build](https://github.com/covid19r/covid19swiss/workflows/build/badge.svg?branch=master)](https://github.com/covid19r/covid19swiss/actions?query=workflow%3Abuild)
[![CRAN\_Status\_Badge](https://www.r-pkg.org/badges/version/covid19swiss)](https://cran.r-project.org/package=covid19swiss)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)
[![GitHub commit](https://img.shields.io/github/last-commit/covid19r/covid19swiss)](https://github.com/covid19r/covid19swiss/commit/master)
<!-- badges: end -->


The covid19swiss R package provides a tidy format dataset of the 2019 Novel Coronavirus COVID-19 (2019-nCoV) pandemic outbreak in Switzerland cantons and the Principality of Liechtenstein (FL). The `covid19swiss` dataset includes the following fields:

* `date` - the timestamp of the case, a `Date` object
* `canton` - the cantons of Switzerland and Principality of Liechtenstein (FL) abbreviation code
* `type` - the type of case
* `cases` - the number of cases corresponding to the date and case type fields
* `gn_a1_code` - a canton index code for merging geometry data from the **rnaturalearth** package


Where the available `type` field includes the following cases:

* `total_tested` - number of tests performed as of the date
* `total_confirmed` - number of positive cases as of the date
* `new_hosp` - new hospitalizations with respect to the previously reported date
* `current_hosp` - number of hospitalized patients as of the current date
* `current_icu` - number of hospitalized patients in ICUs as of the current date
* `current_vent` - number of hospitalized patients requiring ventilation as of the current date
* `total_recovered` - total number of patients recovered as of the current date
* `total_death` - total number of death as of the current date


More information can be found the following vignettes:

* [Introduction to the covid19swiss Dataset](https://covid19r.github.io/covid19swiss/articles/intro_covid19swiss.html)
* [Updating the covid19swiss Dataset](https://covid19r.github.io/covid19swiss/articles/update_the_data.html)
* [Geospatial Visualization of Switzerland Cantons](https://covid19r.github.io/covid19swiss/articles/spatial_dataviz.html) (non CRAN vignette)


Data source: [Specialist Unit for Open Government Data Canton of Zurich
](http://open.zh.ch/internet/justiz_inneres/ogd/de/daten.html), raw data is available on the following [repository](https://github.com/openZH/covid_19). Special thanks for all the people that collaborate and contribute to pull the data from multiple sources and make this data available!




## Installation

You can install the released version of covid19swiss from [CRAN](https://cran.r-project.org/web/packages/covid19swiss/index.html) with:

``` r
install.packages("covid19swiss")
```

Or, install the most recent version from [GitHub](https://github.com/Covid19R/covid19swiss) with:

``` r
# install.packages("devtools")
devtools::install_github("Covid19R/covid19swiss")
```

## Data refresh

The **covid19swiss** package dev version is been updated on a daily bases. The `update_swiss_data` function enables a simple refresh of the installed package datasets with the most updated version on Github:

``` r
library(covid19swiss)

update_swiss_data()
```

More information about updating the data is available on this [vignette](https://covid19r.github.io/covid19swiss/articles/update_the_data.html)

**Note:** must restart the R session after using the `update_swiss_data` function in order to have the updates available


## Usage


```{r}
data(covid19swiss)

head(covid19swiss)
```


### Wide format

```{r}
library(tidyr)

covid19swiss_wide <- covid19swiss %>% 
  pivot_wider(names_from = type, values_from = cases)

head(covid19swiss_wide)

```

## Missing values

The data is collected from multiple resources on the canton level, and not necessarily each data resource provides the same field. Therefore, some fields, such as total recovered or total tested may not be available at this point for some cantons and are marked as missing values (i.e., `NA`). 



## Contributing

Please submit [issues](https://github.com/Covid19R/covid19swiss/issues)
and [pull requests](https://github.com/Covid19R/covid19swiss/pulls)
with any package improvements\!

Please note that the ‘covid19swiss’ project is released with a
[Contributor Code of Conduct](https://github.com/Covid19R/covid19swiss/blob/master/CODE_OF_CONDUCT.md). By contributing to
this project, you agree to abide by its terms.
