---
title: "Introduction to the covid19italy Dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the covid19italy Dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message=FALSE, 
  warning=FALSE,
  comment = "#>"
)
```

```{r setup}

```


The covid19swiss R package provides a tidy format dataset of the 2019 Novel Coronavirus COVID-19 (2019-nCoV) pandemic outbreak in Switzerland cantons and Principality of Liechtenstein (FL). 

### Data structure

The `covid19swiss` dataset includes the following fields:

* `date` - the timestamp of the case, a `Date` object
* `canton` - the cantons of Switzerland and Principality of Liechtenstein (FL) abbreviation code
* `type` - the type of case
* `cases` - the number of cases corresponding to the date and case type fields
* `gn_a1_code` - a canton index code for merging geometry data from the **rnaturalearth** package


Where the available `type` field includes the following cases:

* `total_tested` - number of tests performed as of the date
* `total_confirmed` - number of positive cases as of the date
* `new_hosp` - new hospitalizations with respect to the previously reported date
* `current_hosp` - number of hospitalized patients as of the current date
* `current_icu` - number of hospitalized patients in ICUs as of the current date
* `current_vent` - number of hospitalized patients requiring ventilation as of the current date
* `total_recovered` - total number of patients recovered as of the current date
* `total_death` - total number of death as of the current date


The data organized in a long format:

```{r}
library(covid19swiss)

head(covid19swiss)
```


It is straightforward to transform the data into a wide format with the `pivot_wider` function from the **tidyr** package:

```{r}
library(tidyr)

covid19swiss_wide <- covid19swiss %>% 
  pivot_wider(names_from = type, values_from = cases)

head(covid19swiss_wide)
```


### Query and summarise the data



```{r }
library(dplyr)

covid19swiss %>%
  filter(date == as.Date("2020-04-08"),
         type %in% c("total_confirmed", "total_recovered", "total_death")) %>%
  select(canton, cases, type) %>%
  pivot_wider(names_from = type, values_from = cases) %>%
  arrange(-total_confirmed)
```


